Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-05-17 articles/formalism/article.tex
Added new plot generation scripts: 1. euler angles plots and movie, 2. plot elucidating the radiation and convolution. | Tested the existing codes for conversion to qu2eb and eb2qu and qu2equ&bqu
Minor edits to Euler angles plot/movie. Added appendix in formalism paper deriving the important identity on product of spin spherical harmonics and the radiating and convolution interpretations of the kernel.
generated a proper sign version of the codes, generated the convolution kernel plots and added them to the pape. the kernel plots are updated to have the proper sign.
updated figures, table of c

\usepackage[latin1]{inputenc}
\usepackage{amsmath, amsfonts, amssymb, hyperref, graphicx, multirow, subfigure, bbold}
\hypersetup{colorlinks=True,citecolor=blue}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\renewcommand{\thesubfigure}{(\roman{subfigure})}
%\renewcommand{\thesubfigure}{}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqry}{\begin{eqnarray}}
\newcommand{\eeqry}{\end{eqnarray}}
\newcommand{\beqrys}{\begin{subequations}\begin{eqnarray}}
\newcommand{\eeqrys}{\end{eqnarray}\end{subequations}}
\newcommand{\fqu}{\begin{bmatrix} Q \\ U \end{bmatrix}}
\newcommand{\xpm}{\begin{bmatrix} _{+2}X \\ _{-2}X \end{bmatrix}}
\newcommand{\equ}{\begin{bmatrix} Q \\ U \end{bmatrix}_{E}}
\newcommand{\bqu}{\begin{bmatrix} Q \\ U \end{bmatrix}_{B}}
\newcommand{\eblm}{\begin{bmatrix} a^E \\ a^{B} \end{bmatrix}}
\newcommand{\qutox}{\begin{bmatrix} \mathbb{1} & i\mathbb{1}  \\  \mathbb{1} & -i\mathbb{1}   \end{bmatrix}}
\newcommand{\qutoxd}{\begin{bmatrix} \mathbb{1} & \mathbb{1}  \\  -i\mathbb{1} & i\mathbb{1}   \end{bmatrix}}
\newcommand{\ymat}{\begin{bmatrix} _{+2}Y & 0  \\  0 & _{-2}Y  \end{bmatrix}}
\newcommand{\ymatc}{\begin{bmatrix} _{+2}Y^* & 0  \\  0 & _{-2}Y^*  \end{bmatrix}}
\newcommand{\yzmat}{\begin{bmatrix} _{0}Y & 0  \\  0 & _{0}Y  \end{bmatrix}}
\newcommand{\yzmatc}{\begin{bmatrix} _{0}Y^* & 0  \\  0 & _{0}Y^*  \end{bmatrix}}
\newcommand{\ope}{\hat{O}_{E}}
\newcommand{\opb}{\hat{O}_{B}}
\newcommand{\bmat}{\begin{bmatrix}}
\newcommand{\emat}{\end{bmatrix}}
\newcommand{\mdi}{\mathcal{D}/\mathcal{I}}
\newcommand{\mm}{\mathcal{M}}
\newcommand{\md}{\mathcal{D}}
\newcommand{\mi}{\mathcal{I}}
\def\vp#1{$\bar{P}_{#1}$ }
\def\vs{$\bar{S}~$}
\def\eq#1{{Eq.~(\ref{#1})}}
\def\sec#1{{Sec.~\ref{#1}}}
\def\fig#1{{Fig.~\ref{#1}}}
\def\app#1{{Appendix~\ref{#1}}}
\def\ymat#1{\begin{bmatrix} _{+2}Y_{#1} & 0  \\  0 & _{-2}Y_{#1}  \end{bmatrix}}
\def\ymatc#1{\begin{bmatrix} _{+2}Y^{T*}_{#1} & 0  \\  0 & _{-2}Y^{T*}_{#1}  \end{bmatrix}}
\def\yzmat#1{\begin{bmatrix} _{0}Y_{#1} & 0  \\  0 & _{0}Y_{#1}  \end{bmatrix}}
\def\yzmatc#1{\begin{bmatrix} _{0}Y^*_{#1} & 0  \\  0 & _{0}Y^*_{#1}  \end{bmatrix}}

@@

\graphicspath{ {"../../figures/"//} }

@@

\abstract{
  \noindent 
  \revisit{  We derive full-sky, real-space operators that convert between polarization Stokes $Q$/$U$ parameters to the coordinate-independent, scalar $E$/$B$ modes that are widely used in Cosmic Microwave Background and cosmic shear analysis.  The covolution kernels split naturally into angular and radial parts, and we show explicitly how the spatial extent of the convolution kernel depends on the targeted band-limit.  We show that a arbitrary radial dependence can produce $E$/$B$-like maps and that these are simply filtered versions of the standard $E$/$B$ maps.  This allows us to compute $E$/$B$ maps in real space with a compactly-supported kernel, an approach that can guarantee the avoidance of known foreground regions and can be employed in a massively parallel scheme at high-resolution.  We can compute power spectra using standard techniques, and recover the power spectrum of the sky with a simple window function.  
    %
     We cast the standard CMB polarization analysis operators in a matrix-vector notation which facilitates the derivations and shows that the kernels relate directly to spin-0 $Y_{\ell 2}$ spherical harmonic functions.  This new notation also allows us to derive real space operators which decompose the measured Stokes parameters into their even and odd-parity parts, without ever evaluating the scalar $E$/$B$ fields themselves.  This paper is the first part in a series of papers that explore real-space computation of polarization modes and their applications.
  } 
  \\
  \\
%  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}
}

\maketitle
\begingroup
\let\clearpage\relax
\include{./sections/introduction}
\include{./sections/real_space_formalism}
\include{./sections/kernel_discussion}
\include{./sections/generalizedeb}
\include{./sections/discussion}
\endgroup

>>>>

\usepackage[latin1]{inputenc}
\usepackage{amsmath, amsfonts, amssymb, hyperref, graphicx, multirow, subfigure, bbold}
\hypersetup{colorlinks=True,citecolor=blue}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{caption}
\renewcommand{\thesubfigure}{(\roman{subfigure})}
%\renewcommand{\thesubfigure}{}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqry}{\begin{eqnarray}}
\newcommand{\eeqry}{\end{eqnarray}}
\newcommand{\beqrys}{\begin{subequations}\begin{eqnarray}}
\newcommand{\eeqrys}{\end{eqnarray}\end{subequations}}
\newcommand{\fqu}{\begin{bmatrix} Q \\ U \end{bmatrix}}
\newcommand{\xpm}{\begin{bmatrix} _{+2}X \\ _{-2}X \end{bmatrix}}
\newcommand{\equ}{\begin{bmatrix} Q \\ U \end{bmatrix}_{E}}
\newcommand{\bqu}{\begin{bmatrix} Q \\ U \end{bmatrix}_{B}}
\newcommand{\eblm}{\begin{bmatrix} a^E \\ a^{B} \end{bmatrix}}
\newcommand{\qutox}{\begin{bmatrix} \mathbb{1} & i\mathbb{1}  \\  \mathbb{1} & -i\mathbb{1}   \end{bmatrix}}
\newcommand{\qutoxd}{\begin{bmatrix} \mathbb{1} & \mathbb{1}  \\  -i\mathbb{1} & i\mathbb{1}   \end{bmatrix}}
\newcommand{\ymat}{\begin{bmatrix} _{+2}Y & 0  \\  0 & _{-2}Y  \end{bmatrix}}
\newcommand{\ymatc}{\begin{bmatrix} _{+2}Y^* & 0  \\  0 & _{-2}Y^*  \end{bmatrix}}
\newcommand{\yzmat}{\begin{bmatrix} _{0}Y & 0  \\  0 & _{0}Y  \end{bmatrix}}
\newcommand{\yzmatc}{\begin{bmatrix} _{0}Y^* & 0  \\  0 & _{0}Y^*  \end{bmatrix}}
\newcommand{\ope}{\hat{O}_{E}}
\newcommand{\opb}{\hat{O}_{B}}
\newcommand{\bmat}{\begin{bmatrix}}
\newcommand{\emat}{\end{bmatrix}}
\newcommand{\mdi}{\mathcal{D}/\mathcal{I}}
\newcommand{\mm}{\mathcal{M}}
\newcommand{\md}{\mathcal{D}}
\newcommand{\mi}{\mathcal{I}}
\def\vp#1{$\bar{P}_{#1}$ }
\def\vs{$\bar{S}~$}
\def\eq#1{{Eq.~(\ref{#1})}}
\def\sec#1{{Sec.~\ref{#1}}}
\def\fig#1{{Fig.~\ref{#1}}}
\def\app#1{{Appendix~\ref{#1}}}
\def\ymat#1{\begin{bmatrix} _{+2}Y_{#1} & 0  \\  0 & _{-2}Y_{#1}  \end{bmatrix}}
\def\ymatc#1{\begin{bmatrix} _{+2}Y^{\dag}_{#1} & 0  \\  0 & _{-2}Y^{\dag}_{#1}  \end{bmatrix}}
\def\yzmat#1{\begin{bmatrix} _{0}Y_{#1} & 0  \\  0 & _{0}Y_{#1}  \end{bmatrix}}
\def\yzmatc#1{\begin{bmatrix} _{0}Y^*_{#1} & 0  \\  0 & _{0}Y^*_{#1}  \end{bmatrix}}

@@

\graphicspath{ {"./figures/"//} }

@@

\abstract{
  \noindent 
  \revisit{  We derive full-sky, real-space operators that convert between polarization Stokes $Q$/$U$ parameters to the coordinate-independent, scalar $E$/$B$ modes that are widely used in Cosmic Microwave Background and cosmic shear analysis.  The covolution kernels split naturally into angular and radial parts, and we show explicitly how the spatial extent of the convolution kernel depends on the targeted band-limit.  We show that an arbitrary radial dependence can produce $E$/$B$-like maps and that these are simply filtered versions of the standard $E$/$B$ maps.  This allows us to compute $E$/$B$ maps in real space with a compactly-supported kernel, an approach that can guarantee the avoidance of known foreground regions and can be employed in a massively parallel scheme at high-resolution.  We can compute power spectra using standard techniques, and recover the power spectrum of the sky with a simple window function.  The spatial structure of the real space operator provides some intuition for the $E/B$ structure of polarized, filamentary galactic foregrounds.
    %
     We cast the standard CMB polarization analysis operators in a matrix-vector notation which facilitates the derivations and shows that the kernels relate directly to spin-0 $Y_{\ell 2}$ spherical harmonic functions.  This new notation also allows us to derive real space operators which decompose the measured Stokes parameters into their even and odd-parity parts, without ever evaluating the scalar $E$/$B$ fields themselves.  This paper is the first part in a series of papers that explore real-space computation of polarization modes and their applications.
  } 
  \\
  \\
%  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}
}

\maketitle
\begingroup
\let\clearpage\relax
\include{./sections/introduction}
\include{./sections/real_space_formalism}
\include{./sections/kernel_discussion}
\include{./sections/generalizedeb}
\include{./sections/discussion}
\include{./sections/acknowledgments}
\include{./sections/appendix}
\endgroup

### 2018-05-17 articles/formalism/sections/real_space_formalism.tex
Editing pass and notation cleanup.
Modified section pure E/B section
More details on filament discussion and other minor edits.

Given these transformation properties, the combination ${}_{+2}X(\hat{n}_q)e^{-i2\alpha'_{qe}}$ is invariant under rotations and must be spin-0 by definition:
\beq
\mathcal{R}_{\hat{z}_q}(\phi)[{}_{+2}X(\hat{n}_q)e^{-i2\alpha'_{qe}}] = {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}
\eeq
The real part of the function is constructed by product of functions $(Q\cos{2\alpha}, U\sin{2 \alpha})$ having the same parity and hence the real part must have even parity.  At the same time, the imaginary part of the function is constructed by multiplying functions $(Q\sin{2 \alpha}, U\cos{2\alpha})$ of opposite parity and hence must have an odd parity. Therefore we can make the association that contributions to  $(E+iB)(\hat n_e)$ must be proportional to $ {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}$.

@@

This type of construction can be generalized to transform a field of any spin to a field of any other spin, not just two and zero, and so we can use a similar construction (in the opposite direction) to transform $E/B$ maps back to the Stokes parameters (i.e. transforming spin-0 fields to spin-2).

@@

\paragraph{Convolution kernel.} We can also formulate the real space operator as a convolution operation, where the scalar field at $\hat n_e$ gathers contributions from the Stokes fields.  This is based around the inverse rotation from the previous section (to align the coordinate system at $\hat{n}_e$ with that at $\hat{n}_q$).  The inverse rotation Euler angles relates to the forward rotation Euler angles by the following relations: $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$. Since the kernel only depends on the cosine of the Euler angle $\beta$, it is immune to changes in its sign. The operator equation can be expressed as a function of the Euler angle $\gamma_{eq}$ as follows, 
%
\beq \label{eq:qu2eb_convolution_explicit}
\bmat E_e \\ B_e  \emat =- \sum_{q=1}^{N_{\rm pix}}{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \gamma_{eq}) & -\sin(2\gamma_{eq})\\  \sin(2 \gamma_{eq})  & \cos(2 \gamma_{eq}) \emat  \bmat Q_q \\ U_q  \emat \Delta \Omega\,,
\eeq
%
%where $\Delta \Omega$ denotes the pixel area and all the symbols have their usual meaning. 
\revisit{This version of the real space operator would have naturally emerged had we simplified the the equivalent function $\sum_{\ell m} {}_{0}Y^*_{\ell m}(\hat{n}_e){}_{+2}Y_{\ell m}(\hat{n}_q)$ in \eq{eq:qu2eb_gen_kernel}.} This formulation of the real space operator can be interpreted as integrating at some fixed location $\hat{n}_e$ the $E/B$ mode contribution arising from the Stokes parameters at all location $\hat{n}_q$ on the sphere. This operation can be expressed more concisely as follows,
%
\begin{subequations} \label{q:qu2eb_convolution_concise}
\beqry 
[E + iB](\hat{n}_e) &=& - \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left(\sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} P_{\ell}^{2}(\beta_{eq}) \right) {\Bigg( e^{i2 \gamma_{eq}}   {}_{+2}X (\hat{n}_q) \Bigg)}, \label{eq:qu2eb_physical}\\
&=& \Bigg\lbrace \left[ - \Delta \Omega \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \sqrt{\frac{2 \ell+1}{4 \pi}}Y_{\ell 2}(\beta_{eq},\gamma_{eq}) \right]  \star {}_{+2}X \Bigg\rbrace(\hat{n}_e) \,, \\
&=& \Bigg\lbrace \mathcal{M}_{B} \star {}_{+2}X \Bigg\rbrace(\hat{n}_e) \,, \label{eq:qu2eb_convolution} 
\eeqry
\end{subequations}
%
where $\star$ denotes a convolution and $\mathcal{M}_{B}$ is merely the conjugated function $\mm^*$ when expressed as a function of the Euler angle $\gamma_{eq}$ it can be thought of as an effective instrument beam pointing to the direction $\hat{n}_e$. 
%--------------------------------------------------------

@@

\subsection{Purifying Stokes parameters $Q$/$U$ from $E$/$B$ modes}

@@

The operator $\mathcal{I}$ is Hermitian and is a band limited version of the delta function owing to the identity $\lim_{\ell \rightarrow \infty} \mathcal{I} = \delta(\hat{n}_i - \hat{n}_j)$. For all practical purposes $\mathcal{I}$ acts like an identity operator as is confirmed by the following set of identities: (i) $\mathcal{I} \mathcal{I}=\mathcal{I}$ ; (ii) $\mathcal{D} \mathcal{I}=\mathcal{D}$. $\mathcal{D}$ is a complex but symmetric matrix and $\mathcal{D}^*$ is its inverse in this band limited sense: $\mathcal{D}^* \mathcal{D}=\mathcal{I}$. Using these properties of the operators $\mathcal{I}$ and $\mathcal{D}$, one can verify that the real space operators satisfy the following identities,
%
\begin{subequations}
\beqry
\bar O_E  \bar O_E &=& \bar O_E;\qquad \bar O_B * \bar O_B = \bar O_B \,, \\
\bar O_E \bar O_B &=& 0 \,,\\
\bar O_E + \bar O_B &=& \mathcal{I} \,,
\eeqry
\end{subequations}
%
which are the real space analogues of \eq{eq:har_op_prop}.\footnote{While testing the above stated identities one encounters terms like $\mathcal{D} \mathcal{I}^*,\mathcal{I}^*\mathcal{I}$ and $\mathcal{I}\mathcal{I}^*$ which cannot be simply interpreted but they always occur in pairs with opposite signs that exactly cancel each other.}

>>>>

Given these transformation properties, the combination ${}_{+2}X(\hat{n}_q)e^{-i2\alpha_{qe}}$ is invariant under rotations and must be spin-0 by definition:
\beq
\mathcal{R}_{\hat{z}_q}(\phi)[{}_{+2}X(\hat{n}_q)e^{-i2\alpha'_{qe}}] = {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}
\eeq
\revisit{I'm having some trouble understanding the next 2 sentences.} The real part of the function is constructed by product of functions $(Q\cos{2\alpha}, U\sin{2 \alpha})$ having the same parity and hence the real part must have even parity.  At the same time, the imaginary part of the function is constructed by multiplying functions $(Q\sin{2 \alpha}, U\cos{2\alpha})$ of opposite parity and hence must have an odd parity. Therefore we can make the association that contributions to  $(E+iB)(\hat n_e)$ must be proportional to $ {}_{+2}X(\hat{n}_q) e^{-i2\alpha_{qe}}$.

@@

This type of real-space construction can be generalized to transform a field of any spin to a field of any other spin, not just two and zero, and so we can use a similar construction (in the opposite direction) to transform $E/B$ maps back to the Stokes parameters (i.e. transforming spin-0 fields to spin-2).

@@

\paragraph{Convolution kernel.} We can also formulate the real space operator as a convolution operation, where the scalar field at $\hat n_e$ gathers contributions from the Stokes fields.  This is based around the inverse rotation from the previous section (to align the coordinate system at $\hat{n}_e$ with that at $\hat{n}_q$).  The inverse rotation Euler angles relates to the forward rotation Euler angles by the following relations: $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$. Since the kernel only depends on the cosine of the Euler angle $\beta$, it is immune to changes in its sign. The operator equation can be expressed as a function of the Euler angle $\gamma_{eq}$ as follows, 
%
\beq \label{eq:qu2eb_convolution_explicit}
\bmat E_e \\ B_e  \emat =- \sum_{q=1}^{N_{\rm pix}}{{}_{\mm}f}(\beta_{eq},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \gamma_{eq}) & -\sin(2\gamma_{eq})\\  \sin(2 \gamma_{eq})  & \cos(2 \gamma_{eq}) \emat  \bmat Q_q \\ U_q  \emat \Delta \Omega\,,
\eeq
%
%where $\Delta \Omega$ denotes the pixel area and all the symbols have their usual meaning. 
\revisit{This version of the real space operator would have naturally emerged had we simplified the the equivalent function $\sum_{\ell m} {}_{0}Y^*_{\ell m}(\hat{n}_e){}_{+2}Y_{\ell m}(\hat{n}_q)$ in \eq{eq:qu2eb_gen_kernel}.} This formulation of the real space operator can be interpreted as integrating at some fixed location $\hat{n}_e$ the $E/B$ mode contribution arising from the Stokes parameters at all location $\hat{n}_q$ on the sphere. This operation can be expressed more concisely as follows,
%
\begin{subequations} \label{q:qu2eb_convolution_concise}
\beqry 
[E + iB](\hat{n}_e) &=& - \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left(\sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} P_{\ell}^{2}(\beta_{eq}) \right) {\Bigg( e^{i2 \gamma_{eq}}   {}_{+2}X (\hat{n}_q) \Bigg)}, \label{eq:qu2eb_physical}\\
&=& \Bigg\lbrace \left[ - \Delta \Omega \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \sqrt{\frac{2 \ell+1}{4 \pi}}Y_{\ell 2}(\beta_{eq},\gamma_{eq}) \right]  \star {}_{+2}X \Bigg\rbrace(\hat{n}_e) \,, \\
&=& \Bigg\lbrace \mathcal{M}_{B} \star {}_{+2}X \Bigg\rbrace(\hat{n}_e) \,, \label{eq:qu2eb_convolution} 
\eeqry
\end{subequations}
%
where $\star$ denotes a convolution and $\mathcal{M}_{B}$ is merely the conjugated function $\mm^*$.  When expressed as a function of the Euler angle $\gamma_{eq}$ it can be thought of as an effective instrument beam pointing to the direction $\hat{n}_e$. 
%--------------------------------------------------------

@@

\subsection{Purifying Stokes parameters $Q$/$U$ for $E$/$B$ modes}

@@

The operator $\mathcal{I}$ is Hermitian and is a band limited version of the delta function owing to the identity $\lim_{\ell \rightarrow \infty} \mathcal{I} = \delta(\hat{n}_i - \hat{n}_j)$. For all practical purposes $\mathcal{I}$ acts like an identity operator as is confirmed by the following set of identities: (i) $\mathcal{I} \mathcal{I}=\mathcal{I}$ ; (ii) $\mathcal{D} \mathcal{I}=\mathcal{D}$. $\mathcal{D}$ is a complex but symmetric matrix and $\mathcal{D}^*$ is its inverse in this band limited sense: $\mathcal{D}^* \mathcal{D}=\mathcal{I}$. Using these properties of the operators $\mathcal{I}$ and $\mathcal{D}$, one can verify that the real space operators satisfy the following identities,
%
\begin{subequations}
\beqry
\bar O_E \, \bar O_E &=& \bar O_E;\qquad \bar O_B \, \bar O_B = \bar O_B \,, \\
\bar O_E \, \bar O_B &=& 0 \,,\\
\bar O_E + \bar O_B &=& \mathcal{I} \,,
\eeqry
\end{subequations}
%
which are the real space analogues of \eq{eq:har_op_prop}.\footnote{While testing the above stated identities one encounters terms like $\mathcal{D} \mathcal{I}^*,\mathcal{I}^*\mathcal{I}$ and $\mathcal{I}\mathcal{I}^*$ which cannot be simply interpreted but they always occur in pairs with opposite signs that exactly cancel each other.}  Thus they are orthogonal and idempotent in a band-limited sense.
